---
title: "BCPP Preliminary Analysis"
author: "Melody Owen"
output: html_document
date: "2025-05-28"
---

```{r setup, include = FALSE}
# Package names
packages <- c("tidyverse", "lme4", "MASS", "Matrix", "matrixcalc", "latex2exp",
              "ICC", "nlme", "bindata", "gee", "crt2power", "mosaic", "haven",
              "tmvtnorm", "ggplot2", "latex2exp", "reshape2", "gridExtra",
              "table1")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```


```{r load_data, include = FALSE}
# Loading in original datasets
dat1 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr1_20samp_open.sas7bdat")
dat2 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr2_20samp_open.sas7bdat")
dat3 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr3_20samp_open.sas7bdat")
```


```{r column_names, include = FALSE}
# Create a dataframe for each set of column names
names1 <- tibble(colname = sort(colnames(dat1)), dataset = "dat1")
names2 <- tibble(colname = sort(colnames(dat2)), dataset = "dat2")
names3 <- tibble(colname = sort(colnames(dat3)), dataset = "dat3")

# Bind and pivot to wide
all_names <- bind_rows(names1, names2, names3) %>%
  mutate(present = colname) %>%
  pivot_wider(names_from = dataset, values_from = present) %>%
  arrange(colname)

all_names_filtered <- bind_rows(names1, names2, names3) %>%
  mutate(present = colname) %>%
  pivot_wider(names_from = dataset, values_from = present) %>%
  arrange(colname) %>%
  dplyr::filter(colname %in% c("de_subj_idC", # Subject identifier
                               "community", # Community identifier
                               
                               "random_arm", # Randomization arm (Treatment Assignment)
                               "gender", # Gender
                               
                               # At start of study? Unclear
                               "hiv_status_current", # Current HIV status
                               "circumcised", # Male circumcision
                               
                               # "Coverage Endpoint"
                               "endpoint_coverage_htc", # Coverage endpoint: HIV-tested or diagnosed HIV+
                               "endpoint_coverage_mc", # Coverage endpoint: male circumcision
                               
                               
                               # Outcome
                               "endpoint_seroconvert") # Incidence endpoint: seroconverted by study completion
  )

# Keeping subject ID the same name so I can match the data
dat1_renamed <- dat1 %>% rename_with(~ paste0(., "_dat1"), -de_subj_idC)
dat2_renamed <- dat2 %>% rename_with(~ paste0(., "_dat2"), -de_subj_idC)
dat3_renamed <- dat3 %>% rename_with(~ paste0(., "_dat3"), -de_subj_idC)
```


```{r merging_datasets include = FALSE}
# Circumcision in all three datasets
combinedDat <- full_join(dat1_renamed, dat2_renamed, by = "de_subj_idC") %>%
  full_join(dat3_renamed, by = "de_subj_idC") %>%
  dplyr::select(de_subj_idC, starts_with("random_arm_dat"), 
                contains("community_dat"),
                starts_with("Gender_dat"), hiv_status_current_dat1,
                starts_with("circumcised_dat"), contains("endpoint_coverage_mc"),
                contains("endpoint_coverage_htc"), contains("seroconvert")) %>%
  mutate(random_arm = ifelse(!is.na(random_arm_dat1), random_arm_dat1,
                             ifelse(!is.na(random_arm_dat2), random_arm_dat2,
                                    random_arm_dat3))) %>%
  mutate(gender = ifelse(!is.na(gender_dat1), gender_dat1,
                         ifelse(!is.na(gender_dat2), gender_dat2,
                                gender_dat3))) %>%
  mutate(community = ifelse(!is.na(community_dat1), community_dat1,
                            ifelse(!is.na(community_dat2), community_dat2,
                                   community_dat3))) %>%
  dplyr::select(subject_ID = de_subj_idC, community, gender,
                random_arm, hiv_status_current = hiv_status_current_dat1,
                starts_with("circumcised"),
                contains("endpoint_coverage_mc"), 
                contains("endpoint_coverage_htc"),
                contains("seroconvert")) %>%
  mutate(gender = case_when(gender == "F" ~ "Female",
                            gender == "M" ~ "Male",
                            TRUE ~ NA_character_
  )) %>%
  mutate(circumcised_dat1 = ifelse(gender == "Female",
                                   "Female", circumcised_dat1),
         circumcised_dat2 = ifelse(gender == "Female",
                                   "Female", circumcised_dat2),
         circumcised_dat3 = ifelse(gender == "Female",
                                   "Female", circumcised_dat3)) %>%
  mutate(circumcised_dat2 = ifelse(circumcised_dat1 == "Yes", "Yes",
                                   circumcised_dat2)) %>%
  mutate(circumcised_dat3 = ifelse(circumcised_dat2 == "Yes", "Yes",
                                   ifelse(circumcised_dat1 == "Yes", "Yes",
                                          circumcised_dat3))) %>%
  mutate(across(everything(), ~ na_if(., ""))) %>%
  mutate(across(where(is.character), str_trim)) %>%
  mutate(endpoint_coverage_mc_dat1 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat1),
         endpoint_coverage_mc_dat2 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat2),
         endpoint_coverage_mc_dat3 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat3)) %>%
  mutate(endpoint_coverage_mc = ifelse(circumcised_dat1 == "Yes",
                                       "Began study circumcised", NA)) %>%
  mutate(endpoint_coverage_mc = ifelse(gender == "Female", "Female",
                                       endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat3 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat2 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat1 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & circumcised_dat3 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & circumcised_dat2 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) &
                                         rowSums(across(contains("dat"), ~ replace_na(. == "No", FALSE))) > 0 &
                                         rowSums(across(contains("dat"), ~ replace_na(. == "Yes", FALSE))) == 0,
                                       "No", endpoint_coverage_mc))
```


```{r cleaning_data include = FALSE}
cleanDat <- combinedDat %>%
  dplyr::select(original_subject_id = subject_ID, 
                original_cluster_id = community, 
                random_arm, gender, 
                hiv_status_current,
                endpoint_coverage_htc = endpoint_coverage_htc_dat1, 
                endpoint_coverage_mc, 
                endpoint_seroconvert = endpoint_seroconvert_dat1) %>%
  arrange(original_cluster_id, original_subject_id) %>%
  mutate_if(is.character, list(~na_if(., ""))) %>% # change any blanks to NAs
  rowid_to_column("subject_id") %>%
  group_by(original_cluster_id) %>%
  mutate(cluster_id = cur_group_id()) %>%
  mutate(subj_cluster_id = row_number()) %>%
  ungroup() %>%
  add_count(cluster_id, name = "cluster_size") %>%
  relocate(subject_id, subj_cluster_id, cluster_id, cluster_size) %>%
  ungroup() %>%
  dplyr::select(-original_subject_id, -original_cluster_id) %>%
  mutate(random_arm = ifelse(random_arm == "Standard of Care", "Control",
                             ifelse(random_arm == "Intervention", "Treatment", NA))) %>%
  mutate(endpoint_seroconvert = ifelse(hiv_status_current == "HIV-infected",
                                       "Began study HIV-infected", endpoint_seroconvert))
  
```

# Baseline Characteristics of BCPP Dataset

## Characteristics of Treatment Groups

Below shows Table 1 for the baseline characteristics of the BCPP dataset. All proportions are calculated using this dataset (no exclusions). Then, we exclude any individuals who start the study HIV-infected, refused HIV testing, or had missing HIV testing. 

```{r creating_table1 echo = FALSE}
table1_cleanDat <- table1(~cluster_size + gender + hiv_status_current + 
                            endpoint_coverage_htc + endpoint_coverage_mc + 
                            endpoint_seroconvert | random_arm, data = cleanDat)
table1_cleanDat

paste0("Excluded from the study: n = ", nrow(cleanDat) - nrow(filter(cleanDat, hiv_status_current == "HIV-uninfected")))
```

## Male Circumcision Characteristics

Below shows more characteristics of male circumcision. 

```{r creating_table1 echo = FALSE}
table2_cleanDat <- table1(~ hiv_status_current + 
                            endpoint_coverage_htc + 
                            endpoint_seroconvert | endpoint_coverage_mc, 
                          data = mutate(cleanDat, 
                                        endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc), "Missing", endpoint_coverage_mc)))
table2_cleanDat
```


## Gender Characteristics

Below shows more characteristics of gender. 

```{r creating_table1 echo = FALSE}
table3_cleanDat <- table1(~ hiv_status_current + 
                            endpoint_coverage_htc + 
                            endpoint_seroconvert | gender, 
                          data = mutate(cleanDat, 
                                        gender = ifelse(is.na(gender), "Missing", gender)))
table3_cleanDat
```

## Outcome Characteristics

Below shows more characteristics of HIV status at the end of the study. 

```{r creating_table1 echo = FALSE}
table4_cleanDat <- table1(~ gender + 
                            endpoint_coverage_mc + 
                            endpoint_coverage_htc | endpoint_seroconvert, 
                          data = mutate(cleanDat, 
                                        endpoint_seroconvert = ifelse(is.na(endpoint_seroconvert), "Missing", endpoint_seroconvert)))
table4_cleanDat
```


# Modeling Results

```{r create_model_dat include = FALSE}
modelDat <- cleanDat %>%
  # Calculate proportions first
  mutate(T_k = ifelse(random_arm == "Treatment", 1, 
                      ifelse(random_arm == "Control", 0, NA))) %>%
  # 1 if circumcised before or during study, 0 if no, "Female" if female
  mutate(X1_ik = ifelse(endpoint_coverage_mc == "Yes" | endpoint_coverage_mc == "Began study circumcised", "Yes",
                        ifelse(endpoint_coverage_mc == "No", "No", endpoint_coverage_mc))) %>%
  # 1 if HTC yes, 0 if HTC no
  mutate(X2_ik = ifelse(endpoint_coverage_htc == "Yes", 1,
                        ifelse(endpoint_coverage_htc == "No", 0, NA))) %>%
  # 1 if Male, circumcised, and got HTC | 1 if female and got HTC, 0 otherwise
  mutate(X12_ik = ifelse(gender == "Male" & X1_ik == "Yes" & X2_ik == 1, 1, 
                         ifelse(gender == "Female" & X2_ik == 1, 1, 0))) %>%
  # Gender characteristic, 1 if Male, 0 if female
  mutate(C1_ik = ifelse(gender == "Male", 1, 
                        ifelse(gender == "Female", 0, NA))) %>%
  group_by(cluster_id) %>%
  # Count of all men in each cluster
  mutate(male_count = sum(gender == "Male", na.rm = TRUE)) %>%
  # Count of all male circumcised men in each cluster
  mutate(mc_count = sum(X1_ik == "Yes", na.rm = TRUE)) %>%
  # Count of those who received HTC in each cluster
  mutate(htc_count = sum(X2_ik == 1, na.rm = TRUE)) %>%
  # Count of those who received their full treatment in each cluster
  mutate(full_trt_count = sum(X12_ik == 1, na.rm = TRUE)) %>%
  # Count of those who started study HIV positive in each cluster
  mutate(hiv_infected_count = sum(hiv_status_current == "HIV-infected", na.rm = TRUE)) %>%
  ungroup() %>%
  # Treatment related proportions
  mutate(Z1_k = mc_count/male_count,
         Z2_k = htc_count/cluster_size,
         Z12_k = full_trt_count/cluster_size) %>%
  # Characteristic related proportions
  mutate(hiv_infected_prop = hiv_infected_count/cluster_size) %>%
  mutate(Y_ik = ifelse(endpoint_seroconvert == "Yes", 1,
                       ifelse(endpoint_seroconvert == "No", 0, NA))) %>%
  # Then filter out people who already have HIV or refused testing
  dplyr::filter(hiv_status_current == "HIV-uninfected")
```



