---
title: "BCPP Preliminary Analysis"
author: "Melody Owen"
date: "2025-05-30"
output:
  pdf_document:
    toc_depth: 3
    latex_engine: pdflatex  # or xelatex/lualatex if you're using those
    includes:
      in_header: preamble.tex
geometry: margin=0.5in
---

```{r setup, include = FALSE}
# Package names
packages <- c("tidyverse", "lme4", "MASS", "Matrix", "matrixcalc", "latex2exp",
              "ICC", "nlme", "bindata", "gee", "crt2power", "mosaic", "haven",
              "tmvtnorm", "ggplot2", "latex2exp", "reshape2", "gridExtra",
              "table1", "broom", "xtable", "logistf", "kableExtra", "geepack",
              "elrm", "broom", "dplyr", "tibble", "lme4", "broom.mixed",
              "performance", "mice", "glmnet", "rms")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```


```{r load_data, include = FALSE}
# Loading in original datasets
dat1 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr1_20samp_open.sas7bdat")
dat2 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr2_20samp_open.sas7bdat")
dat3 <- read_sas("~/Desktop/3. Causal Inference Mediation/BCPP data/BCPP data/longyr3_20samp_open.sas7bdat")
```


```{r column_names, include = FALSE}
# Create a dataframe for each set of column names
names1 <- tibble(colname = sort(colnames(dat1)), dataset = "dat1")
names2 <- tibble(colname = sort(colnames(dat2)), dataset = "dat2")
names3 <- tibble(colname = sort(colnames(dat3)), dataset = "dat3")

# Bind and pivot to wide
all_names <- bind_rows(names1, names2, names3) %>%
  mutate(present = colname) %>%
  pivot_wider(names_from = dataset, values_from = present) %>%
  arrange(colname)

all_names_filtered <- bind_rows(names1, names2, names3) %>%
  mutate(present = colname) %>%
  pivot_wider(names_from = dataset, values_from = present) %>%
  arrange(colname) %>%
  dplyr::filter(colname %in% c("de_subj_idC", # Subject identifier
                               "community", # Community identifier
                               
                               "random_arm", # Randomization arm (Treatment Assignment)
                               "gender", # Gender
                               
                               # At start of study? Unclear
                               "hiv_status_current", # Current HIV status
                               "circumcised", # Male circumcision
                               
                               # "Coverage Endpoint"
                               "endpoint_coverage_htc", # Coverage endpoint: HIV-tested or diagnosed HIV+
                               "endpoint_coverage_mc", # Coverage endpoint: male circumcision
                               
                               
                               # Outcome
                               "endpoint_seroconvert") # Incidence endpoint: seroconverted by study completion
  )

# Keeping subject ID the same name so I can match the data
dat1_renamed <- dat1 %>% rename_with(~ paste0(., "_dat1"), -de_subj_idC)
dat2_renamed <- dat2 %>% rename_with(~ paste0(., "_dat2"), -de_subj_idC)
dat3_renamed <- dat3 %>% rename_with(~ paste0(., "_dat3"), -de_subj_idC)
```

```{r echo = FALSE}
combinedDat <- full_join(dat1_renamed, dat2_renamed, by = "de_subj_idC") %>%
  full_join(dat3_renamed, by = "de_subj_idC") %>%
  dplyr::select(de_subj_idC, starts_with("random_arm_dat"), 
                contains("community_dat"),
                starts_with("Gender_dat"), hiv_status_current_dat1,
                starts_with("circumcised_dat"), starts_with("circumcision_days"),
                contains("endpoint_coverage_mc"),
                contains("endpoint_coverage_htc"), contains("seroconvert")) 
```


```{r merging_datasets, include = FALSE}
# Circumcision in all three datasets
combinedDat <- full_join(dat1_renamed, dat2_renamed, by = "de_subj_idC") %>%
  full_join(dat3_renamed, by = "de_subj_idC") %>%
  dplyr::select(de_subj_idC, starts_with("random_arm_dat"), 
                contains("community_dat"),
                starts_with("Gender_dat"), hiv_status_current_dat1,
                starts_with("circumcised_dat"), starts_with("circumcision_days"), 
                contains("endpoint_coverage_mc"),
                contains("endpoint_coverage_htc"), contains("seroconvert")) %>%
  mutate(random_arm = ifelse(!is.na(random_arm_dat1), random_arm_dat1,
                             ifelse(!is.na(random_arm_dat2), random_arm_dat2,
                                    random_arm_dat3))) %>%
  mutate(gender = ifelse(!is.na(gender_dat1), gender_dat1,
                         ifelse(!is.na(gender_dat2), gender_dat2,
                                gender_dat3))) %>%
  mutate(community = ifelse(!is.na(community_dat1), community_dat1,
                            ifelse(!is.na(community_dat2), community_dat2,
                                   community_dat3))) %>%
  dplyr::select(subject_ID = de_subj_idC, community, gender,
                random_arm, hiv_status_current = hiv_status_current_dat1,
                starts_with("circumcised"), starts_with("circumcision_days"),
                contains("endpoint_coverage_mc"), 
                contains("endpoint_coverage_htc"),
                contains("seroconvert")) %>%
  mutate(gender = case_when(gender == "F" ~ "Female",
                            gender == "M" ~ "Male",
                            TRUE ~ NA_character_
  )) %>%
  mutate(circumcised_dat1 = ifelse(gender == "Female",
                                   "Female", circumcised_dat1),
         circumcised_dat2 = ifelse(gender == "Female",
                                   "Female", circumcised_dat2),
         circumcised_dat3 = ifelse(gender == "Female",
                                   "Female", circumcised_dat3)) %>%
  mutate(circumcised_dat2 = ifelse(circumcised_dat1 == "Yes", "Yes",
                                   circumcised_dat2)) %>%
  mutate(circumcised_dat3 = ifelse(circumcised_dat2 == "Yes", "Yes",
                                   ifelse(circumcised_dat1 == "Yes", "Yes",
                                          circumcised_dat3))) %>%
  mutate(across(where(~ !is.numeric(.)), ~ na_if(., ""))) %>%
  mutate(across(where(is.character), str_trim)) %>%
  mutate(endpoint_coverage_mc_dat1 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat1),
         endpoint_coverage_mc_dat2 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat2),
         endpoint_coverage_mc_dat3 = ifelse(gender == "Female", "Female", endpoint_coverage_mc_dat3)) %>%
  mutate(endpoint_coverage_mc = ifelse(circumcised_dat1 == "Yes" & circumcision_days_dat1 < 0,
                                       "Began study circumcised", ifelse(circumcised_dat1 == "Yes" & (circumcision_days_dat1 >= 0 | is.na(circumcision_days_dat1)), "Yes", NA))) %>%
  # mutate(endpoint_coverage_mc = ifelse(circumcised_dat1 == "Yes",
  #                                      "Began study circumcised", NA)) %>%
  mutate(endpoint_coverage_mc = ifelse(gender == "Female", "Female",
                                       endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat3 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat2 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & endpoint_coverage_mc_dat1 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & circumcised_dat3 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) & circumcised_dat2 == "Yes",
                                       "Yes", endpoint_coverage_mc)) %>%
  mutate(endpoint_coverage_mc = ifelse(is.na(endpoint_coverage_mc) &
                                         rowSums(across(contains("dat"), ~ replace_na(. == "No", FALSE))) > 0 &
                                         rowSums(across(contains("dat"), ~ replace_na(. == "Yes", FALSE))) == 0,
                                       "No", endpoint_coverage_mc))
```


```{r cleaning_data, include = FALSE}
# This dataset is a cleaned version of the original dataset with NO ONE EXCLUDED YET
cleanDat <- combinedDat %>%
  dplyr::select(original_subject_id = subject_ID, 
                original_glmmter_id = community, 
                random_arm, gender, 
                hiv_status_current,
                endpoint_coverage_htc = endpoint_coverage_htc_dat1, 
                endpoint_coverage_mc, 
                endpoint_seroconvert = endpoint_seroconvert_dat1) %>%
  mutate(endpoint_coverage_full = ifelse(gender == "Male" & (endpoint_coverage_mc == "Yes" | endpoint_coverage_mc == "Began study circumcised") & endpoint_coverage_htc == "Yes", "Yes", 
                         ifelse(gender == "Female" & endpoint_coverage_htc == "Yes", "Yes", "No"))) %>%
  arrange(original_glmmter_id, original_subject_id) %>%
  mutate_if(is.character, list(~na_if(., ""))) %>% # change any blanks to NAs
  rowid_to_column("subject_id") %>%
  group_by(original_glmmter_id) %>%
  mutate(cluster_id = cur_group_id()) %>%
  mutate(subj_glmmter_id = row_number()) %>%
  ungroup() %>%
  add_count(cluster_id, name = "cluster_size") %>%
  relocate(subject_id, subj_glmmter_id, cluster_id, cluster_size) %>%
  ungroup() %>%
  dplyr::select(-original_subject_id, -original_glmmter_id) %>%
  mutate(random_arm = ifelse(random_arm == "Standard of Care", "Control",
                             ifelse(random_arm == "Intervention", "Treatment", NA))) %>%
  mutate(endpoint_seroconvert = ifelse(hiv_status_current == "HIV-infected",
                                       "Began study HIV-infected", endpoint_seroconvert))
```

```{r create_model_dat, include = FALSE}
# ModelDat is the clean dataset but with people excluded
# Only exclude people who already have HIV or refused testing
# Also people who don't have the outcome
modelDat_full <- cleanDat %>%
  # Calculate proportions first
  mutate(T_k = ifelse(random_arm == "Treatment", 1, 
                      ifelse(random_arm == "Control", 0, NA))) %>%
  # 1 if circumcised before or during study, 0 if no, "Female" if female
  mutate(X1_ik = ifelse(endpoint_coverage_mc == "Yes" | endpoint_coverage_mc == "Began study circumcised", "Yes",
                        ifelse(endpoint_coverage_mc == "No", "No", endpoint_coverage_mc))) %>%
  # 1 if HTC yes, 0 if HTC no
  mutate(X2_ik = ifelse(endpoint_coverage_htc == "Yes", 1,
                        ifelse(endpoint_coverage_htc == "No", 0, NA))) %>%
  # 1 if Male, circumcised, and got HTC | 1 if female and got HTC, 0 otherwise
  mutate(X12_ik = ifelse(gender == "Male" & X1_ik == "Yes" & X2_ik == 1, 1, 
                         ifelse(gender == "Female" & X2_ik == 1, 1, 0))) %>%
  # Gender characteristic, 1 if Male, 0 if female
  mutate(C1_ik = ifelse(gender == "Male", 1, 
                        ifelse(gender == "Female", 0, NA))) %>%
  mutate(X1_ik_vmmc = ifelse(endpoint_coverage_mc == "Yes", "Yes", 
                       ifelse(endpoint_coverage_mc %in% c("Began study circumcised", "No"), "No",
                              ifelse(endpoint_coverage_mc == "Female", "Female", NA)))) %>%
  group_by(cluster_id) %>%
  # Count of all men in each cluster
  mutate(male_count = sum(gender == "Male", na.rm = TRUE)) %>%
  # Count of all male circumcised men in each cluster
  mutate(mc_count = sum(X1_ik == "Yes", na.rm = TRUE)) %>%
  # Count of all male VMMC in each cluster
  mutate(vmmc_count = sum(X1_ik_vmmc == "Yes", na.rm = TRUE)) %>%
  # Count of those who received HTC in each cluster
  mutate(htc_count = sum(X2_ik == 1, na.rm = TRUE)) %>%
  # Count of those who received their full treatment in each cluster
  mutate(full_trt_count = sum(X12_ik == 1, na.rm = TRUE)) %>%
  # Count of those who started study HIV positive in each cluster
  mutate(hiv_infected_count = sum(hiv_status_current == "HIV-infected", na.rm = TRUE)) %>%
  mutate(hiv_refused_testing_count = sum(hiv_status_current == "Refused HIV testing", na.rm = TRUE)) %>%
  ungroup() %>%
  # Treatment related proportions
  mutate(Z1_k = mc_count/male_count, # Prop males circumcised
         Z1_k_vmmc = vmmc_count/male_count, # Prop males VMMC
         Z2_k = htc_count/cluster_size,
         Z12_k = full_trt_count/cluster_size) %>%
  # Characteristic related proportions
  mutate(hiv_infected_prop = hiv_infected_count/cluster_size) %>%
  mutate(hiv_refused_testing_prop = hiv_refused_testing_count/cluster_size) %>%
  mutate(Y_ik = ifelse(endpoint_seroconvert == "Yes", 1,
                       ifelse(endpoint_seroconvert == "No", 0, NA)))

# Then filter out people who already have HIV or refused testing
modelDat <- modelDat_full %>% 
  dplyr::filter(hiv_status_current == "HIV-uninfected") %>%
  dplyr::filter(!is.na(Y_ik))

excludedDat <- modelDat_full %>%
  dplyr::filter(hiv_status_current != "HIV-uninfected" | is.na(hiv_status_current) | is.na(Y_ik))
```


```{r echo = FALSE}
vmmcModelDat <- modelDat %>%
  filter(gender == "Male") %>%
  filter(!is.na(X1_ik_vmmc))
```

# Frequency Tables

There are `r nrow(filter(vmmcModelDat, X1_ik_vmmc == "Yes"))` (`r round(nrow(filter(vmmcModelDat, X1_ik_vmmc == "Yes"))/nrow(vmmcModelDat), 2)*100`%) men who are VMMC, and there are `r nrow(filter(vmmcModelDat, X1_ik == "Yes"))` (`r round(nrow(filter(vmmcModelDat, X1_ik == "Yes"))/nrow(vmmcModelDat), 2)*100`%) men who are circumcised (but not necessarily VMMC). There are `r nrow(filter(modelDat, is.na(X1_ik_vmmc)))` individuals who have a missing value for VMMC status, and these observations are excluded in the analysis.

Frequency tables for when we keep in individuals who got circumcised before the study (they are counted as VMMC = No).

```{r comment = ""}
tally(Y_ik ~ X1_ik_vmmc, data = modelDat) # For all individuals
tally(Y_ik ~ X1_ik_vmmc, data = vmmcModelDat)
tally(Y_ik ~ gender, data = vmmcModelDat)
```

# Modeling Y ~ VMMC

Include men only in the model.

$$logit(Y_{ik}) = \beta_0 + \beta_1 \times \text{VMMC}$$

```{r echo = FALSE}
# Models for Y ~ VMMC

# Model not accounting for clustering
model_y_vmmc <- glm(Y_ik ~ VMMC,
                    family = binomial(link = 'logit'),
                    data = vmmcDat_model)

# Model accounting for clustering using GLMM
model_y_vmmc_glmm <- glmer(Y_ik ~ VMMC + (1|cluster_id), # Uses exchangeable structure
                           data = vmmcDat_model, 
                           family = binomial(link = "logit"))

# Model accounting for clustering using GEE
model_y_vmmc_gee <- geeglm(Y_ik ~ VMMC,
                           family = binomial(link = "logit"), # logit link
                           id = cluster_id, # clustering variable
                           data = filter(vmmcDat_model, !is.na(VMMC)),
                           corstr  = "exchangeable") # working correlation
```

```{r, echo = FALSE}
# GLM Model
model_y_vmmc_summary <- summary(model_y_vmmc) # Save model summary

tidy_y_vmmc <- broom::tidy(model_y_vmmc, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = NA)

# GLMM Model
model_y_vmmc_summary_glmm <- summary(model_y_vmmc_glmm) # Save model summary

tidy_y_vmmc_glmm <- broom.mixed::tidy(model_y_vmmc_glmm, effects = "fixed", 
                  conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLMM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(performance::icc(model_y_vmmc_glmm)$ICC_adjusted[[1]], 3))

# GEE Model
model_y_vmmc_summary_gee <- summary(model_y_vmmc_gee) # Save model summary

tidy_y_vmmc_gee <- broom::tidy(model_y_vmmc_gee, conf.int = TRUE, 
                            exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GEE",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(model_y_vmmc_gee$geese$alpha[[1]], 3))

# Overall Output Table for SpW
y_vmmc_Model_Table <- rbind(tidy_y_vmmc, tidy_y_vmmc_glmm) %>%
  rbind(tidy_y_vmmc_gee)
```

```{r, results = 'asis', echo = FALSE}
print(xtable(y_vmmc_Model_Table, 
             caption = "Model Summary for Y Regressed on VMMC"), 
      type = 'latex', include.rownames = FALSE, comment = FALSE)
```

\newpage

# Modeling Y ~ T + VMMC

Include men only in the model.

$$logit(Y_{ik}) = \beta_0 + \beta_1 \times \text{VMMC} + \beta_2 \times T_k$$

```{r echo = FALSE}
# Models for Y ~ VMMC + T_k

# Model not accounting for clustering
model_y_t_vmmc <- glm(Y_ik ~ VMMC + T_k,
                    family = binomial(link = 'logit'),
                    data = vmmcDat_model)

# Model accounting for clustering using GLMM
model_y_t_vmmc_glmm <- glmer(Y_ik ~ VMMC + T_k + (1|cluster_id), # Uses exchangeable structure
                           data = vmmcDat_model, 
                           family = binomial(link = "logit"))

# Model accounting for clustering using GEE
model_y_t_vmmc_gee <- geeglm(Y_ik ~ VMMC + T_k,
                           family = binomial(link = "logit"), # logit link
                           id = cluster_id, # clustering variable
                           data = filter(vmmcDat_model, !is.na(VMMC)),
                           corstr  = "exchangeable") # working correlation
```

```{r, echo = FALSE}
# GLM Model
model_y_t_vmmc_summary <- summary(model_y_t_vmmc) # Save model summary

tidy_y_t_vmmc <- broom::tidy(model_y_t_vmmc, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = NA)

# GLMM Model
model_y_t_vmmc_summary_glmm <- summary(model_y_t_vmmc_glmm) # Save model summary

tidy_y_t_vmmc_glmm <- broom.mixed::tidy(model_y_t_vmmc_glmm, effects = "fixed", 
                  conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLMM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(performance::icc(model_y_t_vmmc_glmm)$ICC_adjusted[[1]], 3))

# GEE Model
model_y_t_vmmc_summary_gee <- summary(model_y_t_vmmc_gee) # Save model summary

tidy_y_t_vmmc_gee <- broom::tidy(model_y_t_vmmc_gee, conf.int = TRUE, 
                            exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GEE",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = abs(round(model_y_t_vmmc_gee$geese$alpha[[1]], 3)))

# Overall Output Table for SpW
y_t_vmmc_Model_Table <- rbind(tidy_y_t_vmmc, tidy_y_t_vmmc_glmm) %>%
  rbind(tidy_y_t_vmmc_gee)
```

```{r, results = 'asis', echo = FALSE}
print(xtable(y_t_vmmc_Model_Table, 
             caption = "Model Summary for Y Regressed on VMMC + T"), 
      type = 'latex', include.rownames = FALSE, comment = FALSE)
```



# Modeling Y ~ Z_1

Include men only in the model.

$$logit(Y_{ik}) = \beta_0 + \beta_1 \times Z_k^{(1)}$$

```{r echo = FALSE}
# Models for Y ~ Z1

# Model not accounting for clustering
model_y_z1 <- glm(Y_ik ~ Z1_k,
                    family = binomial(link = 'logit'),
                    data = vmmcDat_model)

# Model accounting for clustering using GLMM
model_y_z1_glmm <- glmer(Y_ik ~ Z1_k + (1|cluster_id), # Uses exchangeable structure
                           data = vmmcDat_model, 
                           family = binomial(link = "logit"))

# Model accounting for clustering using GEE
model_y_z1_gee <- geeglm(Y_ik ~ Z1_k,
                           family = binomial(link = "logit"), # logit link
                           id = cluster_id, # clustering variable
                           data = filter(vmmcDat_model, !is.na(VMMC)),
                           corstr  = "exchangeable") # working correlation
```

```{r, echo = FALSE}
# GLM Model
model_y_z1_summary <- summary(model_y_z1) # Save model summary

tidy_y_z1 <- broom::tidy(model_y_z1, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = NA)

# GLMM Model
model_y_z1_summary_glmm <- summary(model_y_z1_glmm) # Save model summary

tidy_y_z1_glmm <- broom.mixed::tidy(model_y_z1_glmm, effects = "fixed", 
                  conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLMM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(performance::icc(model_y_z1_glmm)$ICC_adjusted[[1]], 3))

# GEE Model
model_y_z1_summary_gee <- summary(model_y_z1_gee) # Save model summary

tidy_y_z1_gee <- broom::tidy(model_y_z1_gee, conf.int = TRUE, 
                            exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GEE",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(model_y_z1_gee$geese$alpha[[1]], 3))

# Overall Output Table for SpW
y_z1_Model_Table <- rbind(tidy_y_z1, tidy_y_z1_glmm) %>%
  rbind(tidy_y_z1_gee)
```

```{r, results = 'asis', echo = FALSE}
print(xtable(y_z1_Model_Table, 
             caption = "Model Summary for Y Regressed on Z1"), 
      type = 'latex', include.rownames = FALSE, comment = FALSE)
```

\newpage


# Modeling Y ~ Z_1 + T_k

Include men only in the model.

$$logit(Y_{ik}) = \beta_0 + \beta_1 \times Z_k^{(1)} + \beta_2 \times T_k$$

```{r echo = FALSE}
# Models for Y ~ Z1

# Model not accounting for clustering
model_y_z1_t <- glm(Y_ik ~ Z1_k + T_k,
                    family = binomial(link = 'logit'),
                    data = vmmcDat_model)

# Model accounting for clustering using GLMM
model_y_z1_t_glmm <- glmer(Y_ik ~ Z1_k + T_k + (1|cluster_id), # Uses exchangeable structure
                           data = vmmcDat_model, 
                           family = binomial(link = "logit"))

# Model accounting for clustering using GEE
model_y_z1_t_gee <- geeglm(Y_ik ~ Z1_k + T_k,
                           family = binomial(link = "logit"), # logit link
                           id = cluster_id, # clustering variable
                           data = filter(vmmcDat_model, !is.na(VMMC)),
                           corstr  = "exchangeable") # working correlation
```

```{r, echo = FALSE}
# GLM Model
model_y_z1_t_summary <- summary(model_y_z1_t) # Save model summary

tidy_y_z1_t <- broom::tidy(model_y_z1_t, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = NA)

# GLMM Model
model_y_z1_t_summary_glmm <- summary(model_y_z1_t_glmm) # Save model summary

tidy_y_z1_t_glmm <- broom.mixed::tidy(model_y_z1_t_glmm, effects = "fixed", 
                  conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLMM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(performance::icc(model_y_z1_t_glmm)$ICC_adjusted[[1]], 3))

# GEE Model
model_y_z1_t_summary_gee <- summary(model_y_z1_t_gee) # Save model summary

tidy_y_z1_t_gee <- broom::tidy(model_y_z1_t_gee, conf.int = TRUE, 
                            exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GEE",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(model_y_z1_t_gee$geese$alpha[[1]], 3))

# Overall Output Table for SpW
y_z1_t_Model_Table <- rbind(tidy_y_z1_t, tidy_y_z1_t_glmm) %>%
  rbind(tidy_y_z1_t_gee)
```

```{r, results = 'asis', echo = FALSE}
print(xtable(y_z1_t_Model_Table, 
             caption = "Model Summary for Y Regressed on Z1 + T"), 
      type = 'latex', include.rownames = FALSE, comment = FALSE)
```


# Modeling Y ~ T_k

Include men only in the model.

$$logit(Y_{ik}) = \beta_0 + \beta_1 \times T_k$$

```{r echo = FALSE}
# Models for Y ~ T

# Model not accounting for clustering
model_y_t <- glm(Y_ik ~ T_k,
                    family = binomial(link = 'logit'),
                    data = vmmcDat_model)

# Model accounting for clustering using GLMM
model_y_t_glmm <- glmer(Y_ik ~ T_k + (1|cluster_id), # Uses exchangeable structure
                           data = vmmcDat_model, 
                           family = binomial(link = "logit"))

# Model accounting for clustering using GEE
model_y_t_gee <- geeglm(Y_ik ~ T_k,
                           family = binomial(link = "logit"), # logit link
                           id = cluster_id, # clustering variable
                           data = filter(vmmcDat_model, !is.na(VMMC)),
                           corstr  = "exchangeable") # working correlation
```

```{r, echo = FALSE}
# GLM Model
model_y_t_summary <- summary(model_y_t) # Save model summary

tidy_y_t <- broom::tidy(model_y_t, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = NA)

# GLMM Model
model_y_t_summary_glmm <- summary(model_y_t_glmm) # Save model summary

tidy_y_t_glmm <- broom.mixed::tidy(model_y_t_glmm, effects = "fixed", 
                  conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GLMM",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(performance::icc(model_y_t_glmm)$ICC_adjusted[[1]], 3))

# GEE Model
model_y_t_summary_gee <- summary(model_y_t_gee) # Save model summary

tidy_y_t_gee <- broom::tidy(model_y_t_gee, conf.int = TRUE, 
                            exponentiate = TRUE) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  mutate(Model = "GEE",
         Term = term,
         `OR [95% CI]` = paste0(round(estimate, 3), " [", 
                            round(conf.low, 3), ", ", 
                            round(conf.high, 3), "]"),
         `p-value` = round(p.value, 3)) %>%
  dplyr::select(Model, Term, `OR [95% CI]`, `p-value`) %>%
  dplyr::filter(Term != "(Intercept)") %>%
  mutate(ICC = round(model_y_t_gee$geese$alpha[[1]], 3))

# Overall Output Table for SpW
y_t_Model_Table <- rbind(tidy_y_t, tidy_y_t_glmm) %>%
  rbind(tidy_y_t_gee)
```

```{r, results = 'asis', echo = FALSE}
print(xtable(y_t_Model_Table, 
             caption = "Model Summary for Y Regressed on Z1 + T"), 
      type = 'latex', include.rownames = FALSE, comment = FALSE)
```






